### 2021-06-14 memo.tex
Bockstein square and β2 memo

\def\bZ{\mathbb{Z}}
\def\diag{\mathop{\mathrm{diag}}}

@@

\subsection{Cohomology of $BPSO(4n'_c+2)$}
Recalling that the $\bZ_2$ cohomology read off from \cite{KonoMimura1974}
allowed us to determine the $\bZ$ cohomology by using Bockstein SS
(see \texttt{wzw-memo.pdf}, copy later), we had
\begin{equation}
	\renewcommand{\arraystretch}{1.2}
	\begin{array}{c|cccccccccccccccc}
		d & 0 & 1 & 2 & 3 & 4 & 5 & 6 & \cdots \\
		\hline
		\multicolumn{1}{l|}{H^d(BPSO(4n'_c+2);\bZ)} & \bZ & 0 & 0 & \bZ_4 & \bZ & 0 & \bZ_2 & \cdots\\
		\multicolumn{1}{l|}{H^d(BPSO(4n'_c+2);\bZ_2)} & \bZ_2 & 0 & \bZ_2 & \bZ_2 & \bZ_2 & \bZ_2 & \bZ_2^{\oplus 2} & \cdots\\
		\hline
		\text{generator} \,(\bZ_2) & 1 & - & a_2 & y'(1) & (a_2)^2 & y'(2) & (a_2)^3 & \cdots\\
		&&&&&&& y'(1)^2\\
	\end{array}
\end{equation}
and if we assume the suspension $\bar\theta: H^\ast(PSO(4n'_c+2);\bZ_2) \to H^\ast(BPSO(4n'_c+2);\bZ_2)$
to commute with Steenrod squares (which seems to be true according to Nishimoto-san's notes),
then these elements are supposed to be related as
\begin{equation*}
	\begin{array}{ccl}
		\beta_2 a_2 & = & y'(1)\\
		Sq^2 y'(1) & = & y'(2)\\
		Sq^1 y'(2) & = & y'(1)^2\\
	\end{array}
\end{equation*}
where $\beta_f$ is a higher Bockstein operator, whose image corresponds to $\bZ_{2^f}$ in integral cohomology.

@@

%\subsection{Adams SS}
%The above analysis might be complemented by Adams SS.
%According to our naive guess,
%the module $\widetilde H^\ast(B\left(\tfrac{SO(4n'_c+2)\times SU(2n_f)}{\bZ_2}\right);\bZ_2)_{\leq 5}$ consists of
%\begin{equation}
%	\begin{tikzpicture}[thick,baseline=(m-2-4.south)]
%		\matrix (m) [
%			matrix of math nodes,
%			row sep= 1em,
%			column sep=6em,
%		]{
%			& \colorB{\bullet} \\
%			& \colorB{\bullet} & \colorC{\bullet}\\
%		    & \colorB{\bullet} & \\
%			\colorA{\bullet} && \colorC{\bullet}\\
%			& \colorB{\bullet} & \\
%			\colorA{\bullet} &&\\
%		};
%		\draw[colorA] (m-4-1.center) to [out=200, in=160] (m-6-1.center);
%		%
%		\node[colorA, right=4pt] at (m-6-1) {$a_2$};
%		\node[colorA, right=4pt] at (m-4-1) {$(a_2)^2$};
%		%
%		\draw[colorB] (m-5-2.center) to [out=160, in=200] (m-3-2.center);
%		\draw[colorB] (m-3-2.center) to (m-2-2.center);
%		\draw[colorB] (m-3-2.center) to [out=160, in=200] (m-1-2.center);
%		%
%		\node[colorB, right=4pt] at (m-5-2) {$y'(1)$};
%		\node[colorB, right=4pt] at (m-3-2) {$y'(2)$};
%		\node[colorB, right=4pt] at (m-2-2) {$y'(1)^2$};
%		\node[colorB, right=4pt] at (m-1-2) {$y'(3)$};
%		%
%		\draw[colorC] (m-2-3.center) to [out=200, in=160] (m-4-3.center);
%		%
%		\node[colorC, right=4pt] at (m-4-3) {$c_2$};
%		\node[colorC, right=4pt] at (m-2-3) {$c_3$};
%	\end{tikzpicture}
%\end{equation}
%This leads to the following Adams chart:
%\begin{center}
%	\begin{sseqdata}[
%		name=M2,
%		Adams grading,
%		classes = fill,
%		xrange = {0}{5},
%		yrange = {0}{3},
%		%,x label = {$t-s$}, y label = {$s$}
%	]
%		\tower[colorA](2,0)
%		\tower[colorA](4,1)
%		\class[white](4,0)
%		\tower[colorC](4,0)
%	\end{sseqdata}
%	\printpage[name = M2,page = 2]
%\end{center}

>>>>

\def\bZ{\mathbb{Z}}
\def\fP{\mathfrak{P}}
\def\diag{\mathop{\mathrm{diag}}}

@@

\subsection{Cohomology of $BPSO(4n'_c+2)$}
Recalling that the $\bZ_2$ cohomology read off from \cite{KonoMimura1974}
allowed us to determine the $\bZ$ cohomology by using Bockstein SS
(see \texttt{wzw-memo.pdf}, copy later), we had
\begin{equation}
	\renewcommand{\arraystretch}{1.2}
	\begin{array}{c|cccccccccccccccc}
		d & 0 & 1 & 2 & 3 & 4 & 5 & 6 & \cdots \\
		\hline
		\multicolumn{1}{l|}{H^d(BPSO(4n'_c+2);\bZ)} & \bZ & 0 & 0 & \bZ_4 & \bZ & 0 & \bZ_2 & \cdots\\
		\multicolumn{1}{l|}{H^d(BPSO(4n'_c+2);\bZ_2)} & \bZ_2 & 0 & \bZ_2 & \bZ_2 & \bZ_2 & \bZ_2 & \bZ_2^{\oplus 2} & \cdots\\
		\hline
		\text{generator} \,(\bZ_2) & 1 & - & a_2 & y'(1) & (a_2)^2 & y'(2) & (a_2)^3 & \cdots\\
		&&&&&&& y'(1)^2\\
	\end{array}
\end{equation}
and if we assume the suspension $\bar\theta: H^\ast(PSO(4n'_c+2);\bZ_2) \to H^\ast(BPSO(4n'_c+2);\bZ_2)$
to commute with Steenrod squares (which seems to be true according to Nishimoto-san's notes),
then these elements are supposed to be related as
\begin{equation*}
	\begin{array}{ccl}
		\beta_2 a_2 & = & y'(1)\\
		Sq^2 y'(1) & = & y'(2)\\
		Sq^1 y'(2) & = & y'(1)^2\\
	\end{array}
\end{equation*}
where $\beta_f$ is a higher Bockstein operator (reduced to $\bZ_2$ cohomology as a whole),
whose image corresponds to $\bZ_{2^f}$ in integral cohomology.

@@

\section{Bockstein square for non-closed cochains}
By definition, the variation of the Pontrjagin square $\fP : H^\bullet(-;\bZ_{2m}) \to H^{2\bullet}(-;\bZ_{4m})$ term is
\begin{equation}
	\label{deltaBockstein}
	\begin{array}{ccl}
		\delta \left(\dfrac{1}{4m}\fP(x)\right)
		& = & \dfrac{1}{4m}\cdot \delta \Big(\widetilde x \cup \widetilde x - \widetilde x \cup_1 \delta \widetilde x\Big)\\
		& = & \dfrac{1}{4m}\cdot\left[\Big(\delta \widetilde x \cup \widetilde x + \widetilde x \cup \delta \widetilde x\Big)
				- \Big(\widetilde x \cup \delta \widetilde x - \delta \widetilde x \cup \widetilde x + \delta \widetilde x \cup_1 \delta \widetilde x\Big)\right]\vspace{2mm}\\
		& = & \dfrac{1}{4m}\cdot\Big[2\cdot \delta \widetilde x \cup \widetilde x - \delta \widetilde x \cup_1 \delta \widetilde x\Big],\\
	\end{array}
\end{equation}
and thus if $x$ were a $\bZ_{2m}$-cocycle $\delta x = 0$ (mod $2m$)
and its lift $\widetilde x$ were $\delta \widetilde x = 0$ or $2m$ (mod $4m$),
then the right hand side would be zero (mod 1) and $\fP(x)$ would be a $\bZ_{4m}$-cocycle (i.e.~$0$ mod 4) as desired.
However, when $x$ is not a cocycle but merely a cochain, this term cannot be defined in the first place.

This can be saved somewhat miraculously as follows.
A short exact sequence
\begin{equation*}
	0
	\to
	\bZ_{2^f}
	\overset{\times 2^{f}}{\longrightarrow}
	\bZ_{2^{2f}}
	\overset{p}{\longrightarrow}
	\bZ_{2^f}
	\to
	0
\end{equation*}
buys us cohomology operations called higher Bockstein $\beta_f: H^{\bullet}(-;\bZ_{2^f}) \to H^{\bullet+1}(-;\bZ_{2^f})$,
and for the element $x \in C^{\bullet}(-;\bZ_{2^f})$ one has
\begin{equation*}
	\delta\widetilde x
	=
	2^f \beta_f (x).
\end{equation*}
Now, let us consider the case of odd $n_c$.
Here, the cochain $w_2(c) \in C^2(SO(4n'_c+2)\times SU(2n_f); \bZ_2)$
can be thought of as a mod-2 reduction of $\widetilde w_2(c)\in C^2(SO(4n'_c+2)\times SU(2n_f); \bZ_4)$.
Dividing the $SO\times SU$ by $\bZ_2$, these become non-closed
\begin{equation*}
	\delta \widetilde w_2(c) = 2 \beta v_2(c)
\end{equation*}
where $v_2(c)=a_2\in C^2\left(\tfrac{SO(4n'_c+2)\times SU(2n_f)}{\bZ_2}; \bZ_2\right)$.
%\begin{equation*}
%	\begin{array}{ccccccccc}
%		&& w_2 && \widetilde w_2 & \longmapsto & v_2\\
%		&& \rotatebox[]{270}{$\in$} && \rotatebox[]{270}{$\in$} && \rotatebox[]{270}{$\in$}\\
%		0
%		& \longrightarrow &
%		C^2(-;\bZ_2)
%		& \overset{i^\ast}{\longrightarrow} &
%		C^2(-;\bZ_4)
%		& \overset{p^\ast}{\longrightarrow} &
%		C^2(-;\bZ_2)
%		& \longrightarrow &
%		0\vspace{1mm}\\
%		&& \downarrow\,\delta && \downarrow\,\delta && \downarrow\vspace{1mm}\\
%		0
%		& \longrightarrow  &
%		C^3(-;\bZ_2)
%		& \longrightarrow &
%		C^3(-;\bZ_4)
%		& \longrightarrow &
%		C^3(-;\bZ_2)
%		& \longrightarrow &
%		0\\
%		&& \rotatebox[]{90}{$\in$} && \rotatebox[]{90}{$\in$}\\
%		&& \delta w_2 && \delta \widetilde w_2
%	\end{array}
%\end{equation*}
Since this implies $\delta \widetilde w_2(c) = 0$ or $2$ mod 4
and furthermore $\delta (2\widetilde w_2(c)) = 0$ mod 4,
one can safely define the Pontrjagin square $\tfrac{1}{8}\fP\big(2\widetilde w_2(c)\big)$.\footnote{
	Be careful that this factor $2$ here is not the usual map sending $\{0,1\}=\bZ_2$ to $\{0,2\}\subset \bZ_4$.
	This time we are \textit{really} multiplying by $2$.
}
Note that this can naively be regarded as $\tfrac{1}{2}\fP\big(w_2(c)\big)$ at the integral cochain level.
While the second term in the last line of \eqref{deltaBockstein} together with the overall factor takes value in
$\tfrac{(2m)^2}{4m}\bZ = m\bZ = 2\bZ$ and dividing by two does not cause any trouble,
the first term does as it takes value in $\tfrac{2\cdot 2m}{4m}\bZ = \bZ$.

Let us take a closer look at the latter.
The long exact sequence of cochain groups implies that $2\widetilde w_2(c) = v_2(c) = a_2$ [YL:make it more precise later].
Then the term of interest is
\begin{equation*}
	\dfrac{1}{8}\cdot 2\cdot 4\beta_2(a_2) \cup a_2.
\end{equation*}
Therefore, the anomalous variation of the Pontrjagin square term seems to result in
\begin{equation*}
	\delta\left(
		\dfrac{1}{4}\fP(w_2(c))
	\right)
	=
	\dfrac{1}{2}a_2 \beta_2 a_2.
\end{equation*}
